---
title: "The fashionability of Flu Shots"
author: "Rachel Donahue"
date: "12/02/2021"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
library(tidyverse)
library(openxlsx)
library(dplyr)
library(usmap)
library(lme4)

unzip("2013-zip.zip")
thirteen<-read.csv("2013.csv")
unzip("2014-zip.zip")
fourteen<-read.csv("2014.csv")
unzip("2015-zip.zip")
fifteen<-read.csv("2015.csv")
unzip("2016-zip.zip")
sixteen<-read.csv("2016.csv")
unzip("2017-zip.zip")
seventeen<-read.csv("2017.csv")
unzip("2018-zip.zip")
eighteen<-read.csv("2018.csv")
unzip("2019-zip.zip")
nineteen<-read.csv("2019.csv")

nineteen$year<-"2019"
eighteen$year<-"2018"
seventeen$year<-"2017"
sixteen$year<-"2016"
fifteen$year<-"2015"
fourteen$year<-"2014"
thirteen$year<-"2013"

data<-rbind(nineteen, eighteen, seventeen, sixteen, fifteen, fourteen, thirteen)
```



## Abstract

Medicare is one of the largest drivers of the healthcare system in the United States with over 62.6 million people enrolled in the program as of 2020 with a projected increase of 50% over the next decade. 

## Introduction



## Method

### Data Cleaning and Processing
The data that is being used to answer this question comes from the Medicare Physician & Other Practicioners- by Provider and Service public use files provided by (CMS) website. This is a fairly large dataset that contains all of the services that providers have submitted claims for through out the year for Original Medicare Part B beneficiaries, broken out by National Provider Identifier (NPI) number and HCPCS code. To subset down to flu shot related charges, I clicked on each of the individual years available (2013-2019), applied the filter button, selected the columns HCPCS_Desc for charges containing the words "flu" and "vaccine".I additionally applied filtering for Rndrng_Prvdr_Cntry="US".


```{r include=FALSE}
dataframe<-data[, c(1:2, 6, 7, 11, 17, 22:23, 26, 29:30)]
names(dataframe)<- c("NPI", "Name", "Gender", "EntityFlag", "State", "PrvdrType", "InOfficeFlag", "Benes", "TotalCharges", "MedicarePaymentAmount", "Year")

dataframe$ThirdPartypaymentamount<-dataframe$TotalCharges-dataframe$MedicarePaymentAmount

flushots<-dataframe %>% filter(State!="AA"&State!="AE"&State!="XX"&State!="AP"&State!="MP"&State!="VI"&State!="GU"&State!="PR")

```
### Exploratory Data Analysis


### Model Fitting

## Result

## Model Coefficients

### Model Validation

## Discussion

## Citation

A Dozen Facts About Medicare Advantage in 2020. (2021, January 13). KFF. https://www.kff.org/medicare/issue-brief/a-dozen-facts-about-medicare-advantage-in-2020/



## Appendix
### More EDA